---
title: "How to Automate Summary Stats in Markdown"
author: "Zach Fogel and ChatGPT"
date: "2024-04-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# These are also helpful
# %btwn% acts like %in%; so 1 %btwn% c(0,2) would return TRUE
# %!in% is the opposite of %in%; 1 %!in% 2:4 returns FALSE
source('R/operator_btwn.R')
source('R/operator_!in.R')
```

## R Markdown

```{r load-data}
# load model
model <- readRDS('Input/wlv-top-model.RData')
# load occupancy data (also includes explanatory variables)
occ <- readRDS('Input/wlv-occ.RData') %>% filter(buffer == 2000)

# Tidy model outputs, add confidence intervals
maketidymod <- function(model) {
  broom::tidy(model) %>%
  mutate(confintlow = estimate - 1.96*std.error,
         confinthigh = estimate + 1.96*std.error) %>% 
  rowwise() %>% 
  # calculate different measures of significance
  mutate(sigCI = !(0 %btwn% c(confintlow, confinthigh)),
         sig.p = p.value < 0.05)
}

printstats <- function(modelsumm) {
  b <- modelsumm$estimate %>% round(3)
  z <- modelsumm$statistic %>% round(3)
  p <- modelsumm$p.value %>% round(3)
  ci1 <- modelsumm$confintlow %>% round(3)
  ci2 <- modelsumm$confinthigh %>% round(3)
  # \\beta outputs as the letter beta
  # %.3f means to use 3 decimal places
  # you can change the values that are included
  message <- sprintf("\\(\\beta = %.3f\\), \\(z = %.3f\\), \\(p = %.3f\\), 95%% CI = [%.3f, %.3f]", b, z, p, ci1, ci2)
  return(message)
}

modelsumm <- maketidymod(model)
stats <- lapply(modelsumm$term, function(x) {
  tmp <- filter(modelsumm, term == x)
  data.frame(term = x, stats = printstats(tmp))
}) %>% bind_rows()
```

For wolverines, the top model included linear disturbance and elevation at a `r model$buffer` m buffer: \[\text{Pres} = \beta_0 + \beta_1(\text{lin}) + \beta_2(\text{elev.sd}) + \beta_3(\text{elev.med}) + \beta_4(\text{days})\]. Linear disturbance (`r filter(stats, term == 'lin')[,2]`) and elevation SD (`r filter(stats, term == 'elev.sd')[,2]`) had signicant effects on wolverine presence.

Chatgpt can probably explain this better, but basically:

- The "\\[ [equation] \\]" means that it's going to be a latex equation block (you probably knew that)
- `r` means "this is going to be an R chunk"
- "filter(stats, term == 'lin')[,2]" is just referring back to the stats object. 